---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%",
  fig.align = "center"
)
```

# classbound

<!-- badges: start -->
[![R-CMD-check](https://github.com/natydasilva/classbound/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/natydasilva/classbound/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/natydasilva/classbound/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/natydasilva/classbound/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

The **classbound** package provides tools to explore, visualize, and analyze classification decision boundaries in R. It offers a unified interface for fitting machine learning models and rendering 2D visualizations of their prediction surfaces.

## Installation

You can install the development version of classbound from GitHub with:

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("natydasilva/classbound")
```

## Quick Start

### Visualizing a Decision Boundary

Here is a basic example of how to fit a model and visualize its decision boundary:

```{r visualisation, message=FALSE, warning=FALSE}
library(classbound)
library(palmerpenguins)

# 1. Load and prepare sample dataset
penguins <- na.omit(penguins[, -c(2, 7, 8)])

# 2. Fit a decision tree (rpart) predicting species
model <- fit_model(
  data = penguins,
  formula = species ~ bill_length_mm + bill_depth_mm,
  classifier = rpart::rpart,
  interface = "formula"
)

# 3. Compute the boundary
feature_range <- list(
  bill_length_mm = c(30.0, 60.0),
  bill_depth_mm = c(10.0, 25.0)
)
model <- boundary_compute(model, range = feature_range, resolution = 100)

# 4. Plot the decision boundary
plot_boundary(
  model = model,
  obs_data = penguins,
  x_col = "bill_length_mm",
  y_col = "bill_depth_mm",
  true_label = "species"
)
```

### Native Confidence Mapping (Gradients)

If the specified classifier returns prediction probabilities (such as Random Forests or Logistic Regression), `classbound` automatically extracts these probabilities and maps them to the visual transparency (`alpha`) of the decision regions. Deep, saturated colors indicate high model confidence, while pale, faded regions indicate model uncertainty near the boundary edges. If a classifier does not return probabilities (e.g., standard SVMs), `classbound` gracefully falls back to drawing solid colors separated by sharp boundaries.

### High-Dimensional Depth Fading (Projections)

When visualizing models with >2 dimensions via projection (using `boundary_compute(..., projection = proj)`), `plot_boundary()` applies mathematically rigorous forward projection to any overlaid training observations.

To prevent visual crowding and provide a sense of spatial depth, it automatically calculates the orthogonal distance from each original high-dimensional point to the 2D projection plane. Points lying exactly on the plane are plotted entirely opaque (`alpha = 1.0`), while points further away in high-dimensional space gradually fade into the background (`alpha = 0.2`). This provides instant visual context on whether a point is genuinely near the boundary or simply an artifact of the viewing angle.

### The All-In-One Wrapper

For convenience, you can perform fitting, boundary computation, and plotting in a single step using the `classbound()` wrapper:

```{r classbound-wrapper, eval=FALSE}
classbound(
  data = penguins,
  formula = species ~ bill_length_mm + bill_depth_mm,
  classifier = rpart::rpart,
  resolution = 100
)
```

### Using Different Classifiers

Because `classbound` uses standard R evaluation, you can easily plug in other models. The package supports standard classifiers out of the box, and provides a `predfun` argument for models that return complex objects (like lists).

```{r classifiers, eval=FALSE}
# Standard models (e.g., SVM) work automatically
model_svm <- classbound(
  data = penguins,
  formula = species ~ bill_length_mm + bill_depth_mm,
  classifier = e1071::svm
)

# Models returning lists (e.g., MASS::qda) use the `predfun` fallback mechanism
model_qda <- classbound(
  data = penguins,
  formula = species ~ bill_length_mm + bill_depth_mm,
  classifier = MASS::qda,
  predfun = function(model, newdata, ...) {
    predict(model, newdata, ...)$class
  }
)
```

### Interactive Exploration

The package also includes a Shiny app to interactively explore models and boundaries. You can launch it using:

```{r eval=FALSE}
explorapp()
```
