---
title: "Custom Predictions and Adapters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Predictions and Adapters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`classbound` is designed to be as simple as possible when dealing with different classifiers. Our prediction philosophy is:

1. **Native R First:** Standard R prediction interfaces work automatically through our default path.
2. **`predfun` Fallback Mechanism:** Models with non-standard documented prediction APIs can be handled directly via the `predfun` argument.
3. **Minimal Built-ins:** Dedicated S3 adapters are only added for officially supported classifiers whose APIs genuinely require special handling.

## 1. Standard Classifiers (No extra code needed)

If your model's `predict()` method returns a vector or factor of classes natively, you don't need to do anything. `classbound` will handle it automatically using R's default dispatch mechanism.

```{r setup_standard, eval=FALSE}
# e1071::svm works out of the box
my_model <- classbound(
  data = my_data,
  formula = y ~ x1 + x2,
  classifier = e1071::svm
)
```

## 2. Non-Standard Models (Using `predfun`)

If your model returns a complex object (like a list, data frame, or matrix), the default path will stop with an error, prompting you to provide a `predfun`. The `predfun` acts as a fallback mechanism, allowing you to explicitly extract the classes without writing an S3 adapter.

```{r setup_predfun, eval=FALSE}
# MASS::qda returns a list, so we supply a predfun
my_model <- classbound(
  data = my_data,
  formula = y ~ x1 + x2,
  classifier = MASS::qda,
  predfun = function(model, newdata, ...) {
    predict(model, newdata, ...)$class
  }
)
```

## 3. Officially Supported Classifiers (Built-in Adapters)

We maintain a very small set of built-in S3 adapters (e.g., `rpart`, `randomForest`, `PPforest`) because their documented APIs require model-specific handling (such as passing `type = "class"` or muffling specific warnings). If you use these classifiers, `classbound` will automatically route them to their built-in adapter. You do not need to provide a `predfun`.

```{r setup_builtin, eval=FALSE}
# PPforest requires model-specific handling for prediction warnings & list outputs,
# but classbound provides a built-in adapter for it natively.
my_model <- classbound(
  data = my_data,
  formula = y ~ x1 + x2,
  classifier = PPforest::PPforest
)
```

### Do I Need to Write a Custom S3 Adapter?

Usually, no! You should only write a custom `predict_adapter` if you are developing an extension package for `classbound` and want to officially support a new complex classifier. Otherwise, simply passing a `predfun` is the easiest and recommended approach.
