---
title: "Custom Predictions and Adapters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Predictions and Adapters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`classbound` is designed to work with the widest possible range of R classifiers.
This vignette explains how prediction routing works and how to handle classifiers
whose APIs do not fit the default path.

## The prediction philosophy

```
Standard classifier (predict returns factor/vector)
    → handled automatically via predict_adapter.default()

Non-standard classifier (predict returns a list or complex object)
    → provide predfun to extract class labels

Officially supported classifiers (rpart, randomForest, PPtree, ppforest2)
    → handled by built-in S3 adapters with full probability support
```

## 1. Standard classifiers — no extra work needed

If a classifier's `predict()` method returns a vector or factor of class labels directly,
`classbound` handles it automatically. No configuration is needed.

```{r standard, eval=FALSE}
library(classbound)
library(palmerpenguins)
penguins <- na.omit(palmerpenguins::penguins[
  ,
  c("species", "bill_length_mm", "bill_depth_mm")
])

# e1071::svm returns a factor of class labels — works out of the box
classbound(penguins, species ~ bill_length_mm + bill_depth_mm, e1071::svm)
```

## 2. Non-standard models — using `predfun`

Some classifiers return a list, data frame, or other complex object from `predict()`.
The default path will stop with an informative error message suggesting you provide
a `predfun`. The `predfun` receives the fitted model and new data, and must return
either a factor/vector of class labels, or a list with `$class` and `$probs`.

```{r predfun, eval=FALSE}
# MASS::qda returns list($class, $posterior, $x) — extract $class
classbound(
  penguins,
  species ~ bill_length_mm + bill_depth_mm,
  MASS::qda,
  predfun = function(model, newdata, ...) predict(model, newdata, ...)$class
)

# MASS::lda — same approach
classbound(
  penguins,
  species ~ bill_length_mm + bill_depth_mm,
  MASS::lda,
  predfun = function(model, newdata, ...) predict(model, newdata, ...)$class
)

# Return probabilities as well (enables gradient visualization)
classbound(
  penguins,
  species ~ bill_length_mm + bill_depth_mm,
  MASS::lda,
  predfun = function(model, newdata, ...) {
    out <- predict(model, newdata, ...)
    list(class = out$class, probs = out$posterior)
  }
)
```

The `predfun` argument is available in `classbound()`, `fit_model()` (via `boundary_compute()`),
and `predict_model()`.

## 3. Officially supported classifiers — built-in adapters

`classbound` maintains a small set of built-in S3 adapters for classifiers whose
APIs require model-specific handling to extract both class labels and probabilities:

| Classifier | Adapter | Probabilities |
|---|---|---|
| `rpart::rpart` | `predict_adapter.rpart` | Yes |
| `randomForest::randomForest` | `predict_adapter.randomForest` | Yes |
| `PPtreeViz::PPTreeclass` | `predict_adapter.PPtreeclass` | No |
| `PPtreeExt::PPtreeExtclass` | `predict_adapter.PPtreeExtclass` | No |
| `ppforest2::pprf` | `predict_adapter.pprf_classification` | Yes |

These adapters are invoked automatically when the classifier object belongs to the
corresponding S3 class. No `predfun` is needed.

## The adapter contract

Every prediction path must produce a list with exactly two elements:

```r
list(
  class = factor(...),  # vector of predicted class labels
  probs = matrix(...)   # n x K probability matrix, or NULL
)
```

`probs` must be `NULL` for classifiers that do not provide probability estimates.
`classbound` handles `NULL` probabilities gracefully: the boundary plot renders with
flat (non-gradient) colored regions instead of a probability surface.

## 4. Writing a custom S3 adapter

Custom S3 adapters are only needed if you are building an extension package for
`classbound` and want to officially support a complex classifier without requiring
users to write `predfun` every time.

For most users, a `predfun` is sufficient and far simpler.

```{r custom_adapter, eval=FALSE}
# Example: custom adapter for a hypothetical classifier "myModel"
predict_adapter.myModel <- function(model, newdata, ...) {
  raw <- predict(model, newdata, type = "response")
  list(
    class = factor(raw$labels),
    probs = as.matrix(raw$probabilities)
  )
}
```

Define the method in your package's namespace and it will be dispatched automatically
whenever `classbound` encounters a model object of class `"myModel"`.
