---
title: "Custom Predict Adapters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Predict Adapters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Do I Need a Custom Adapter?

Most standard R classifiers return a vector or factor of predictions when `predict()` is called. `classbound` provides a "smart default" `predict_adapter.default()` that handles these out of the box. You **only** need to write a custom adapter if your model returns a complex object (like a list, matrix, or custom class) that the default generic cannot handle.

## Adapter Contract

If you do need a custom adapter, you must implement an S3 method for `predict_adapter`.
The function must return a list with two elements:
- `class`: A factor vector of predicted class labels.
- `probs`: A probability matrix, or strictly `NULL` if the classifier does not support probabilities. Downstream pipeline functions handle `probs = NULL` natively.

## Example: Custom S3 Adapter

Here is a skeleton for a custom classifier that returns class `"mymodel"`:

```{r setup, eval=FALSE}
# 1. Define the prediction logic for the native model class
predict_adapter.mymodel <- function(model, newdata, ...) {
  # Get predictions from the custom package (assuming it returns a list)
  raw_preds <- mypackage::predict(model, newdata, ...)
  
  # Extract the class vector
  pred_class <- raw_preds$class
  
  # Extract the probs matrix (or set to NULL if unsupported)
  probs <- raw_preds$probabilities

  list(class = as.factor(pred_class), probs = probs)
}

# 2. Use it natively in the pipeline
# my_model <- fit_model(data, formula, classifier = mypackage::myfit)
```
