---
title: "Interactive Exploration with Explorapp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive Exploration with Explorapp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Launching Explorapp

`explorapp()` opens the interactive Shiny application for visual exploration of
classification decision boundaries. Launch it from the R console:

```{r launch, eval=FALSE}
library(classbound)

# Empty app (start with simulation or drawing)
explorapp()

# Pre-loaded dataset
library(palmerpenguins)
penguins <- na.omit(palmerpenguins::penguins)
explorapp(data = penguins, target_col = "species")
```

When `data` and `target_col` are provided, **Import Data** mode is selected automatically.

---

## Data modes

The sidebar's **Data Mode** panel controls which dataset the models are trained on.

### Import Data

Available when you call `explorapp(data = ..., target_col = ...)`. The provided data
frame is loaded directly. Use **Data Preview** to inspect the loaded rows.

For datasets with more than two features, a **2D Slice** or **Projection** tab
appears in the visualization area (see below).

### Simulate Data

Generate synthetic data from one of two engines:

**Multivariate Normal (MVN)**: You specify the number of classes, and the app
generates class-specific mean and covariance sliders. Click **Generate Data** to
draw a new dataset. A separate independent test set is also generated with
exactly 30% of the training sample size. This is not a 70/30 split; it is
an additional independently drawn evaluation sample using the same class parameters.

**MixSim**: Generates more complex overlapping Gaussian mixture data. Specify the
number of classes `K`, number of dimensions `p`, and maximum pairwise overlap
`MaxOmega`. Smaller overlap values create more clearly separated classes.

Both engines support a **Random Seed** for reproducibility and a **Background Noise**
slider for adding uniformly distributed contamination points.

### Draw Data

Draw your own classification dataset by clicking or brushing directly on the plot.
Choose from two interaction modes:

**Draw Point**: Each click adds one observation at the clicked location. Dragging does
not create a stroke; only the location where you release the mouse counts.

**Draw Cluster**: Dragging creates a cluster of observations along the drawn path.
The **Point Density** and **Brush Size** controls determine how many points are added
per path segment.

Use **Undo Last** to remove the most recently added batch of points. **Clear Canvas**
removes all drawn points.

**Clone to Draw Canvas** copies a supported 2-feature dataset (from Import or Simulate)
onto the Draw Canvas so you can continue editing it. High-dimensional datasets cannot
be cloned because the Draw Canvas is a 2-feature interface.

---

## Model configuration

The **Model Configuration** panel lists all available classifiers. Select one or more
to compare simultaneously. Each classifier shows its own tuning parameters when selected.

Supported classifiers include:
- **rpart** (decision tree): complexity parameter `cp`
- **randomForest**: number of trees, `mtry`
- **PPtreeViz** / **PPtreeExtclass** / **PPtreeExt_split**: projection pursuit trees
- **ppforest2**: projection pursuit random forest
- **Tidymodels presets** (if the `parsnip` package is installed): Decision Tree, Random Forest, SVM, Neural Net, PP Forest

**Grid Resolution** controls the density of the prediction grid: higher values produce
smoother boundaries at the cost of computation time. 100 is a good default.

---

## Visualization modes

### Navigate

The default interaction mode. Use Ctrl + mouse wheel to zoom in/out, or draw a brush
(drag) on the plot to select a region and zoom into it. Double-click to reset the view.

### 2D Slice

For datasets with more than two features, the **2D Slice** tab lets you choose which
two features to display on the axes. All other numeric features are fixed at their
training-set median and categorical features at their mode.

The slice shows the decision boundary at a specific cross-section. Changing the selected
features reveals different slices through the multi-dimensional boundary.

### Projection

The **Projection** tab uses a PCA-derived basis to project the high-dimensional feature
space into two axes. Unlike a slice, the projection combines information from all features
simultaneously.

When a projection is active, training observations are depth-faded: points closer to
the projection plane appear more opaque, while points further away fade toward
transparency.

---

## Probability surface

The **Probability Surface** option (in Visual Settings) shades decision regions by the
model's predicted class probability: deep colors indicate high confidence and faded
colors indicate uncertainty near the boundary.

**Probability surface is only available when the selected classifier provides class
probabilities.** Classifiers such as standard SVMs and PPtree models do not provide
probabilities and always produce flat colored regions regardless of this setting.
The app displays: *"Probability surface unavailable: the selected model does not provide
class probabilities."*

---

## Outlier injection

The **Outlier Injection** panel adds extreme observations to the training data to test
how classifiers respond to unusual points.

- **Outlier Class**: the class label assigned to the injected outlier, or *Random* for
  a randomly chosen class.
- **Outlier Magnitude**: controls how far the outlier is placed from its class distribution.
- **Number of Outliers**: how many outliers to inject at once.
- **Highlight Outliers (Diamonds)**: renders injected outliers with a distinct diamond shape.

Injected outliers become part of the training data. When models are refitted, the outliers
influence the fitted boundary. For simulated data, outliers are added to the training set
and do not affect the independently generated test set.

Use **Clear** to remove all injected outliers without clearing the rest of the dataset.

---

## Performance metrics

The **Training Performance Metrics** table at the bottom of the main panel shows
accuracy, precision, recall, and F1 for each fitted model on the training data.

The **Test Error** column reports error on the independently generated test dataset
(for Simulate Data mode). This test set is generated fresh using the same simulation
parameters as the training data, with approximately 30% of the training sample size.
It is not derived by splitting the training data.

For **Import Data** and **Draw Data** modes, no automatic test data is fabricated.
Test Error is not reported for these modes.

---

## Export

Click **Export Results...** to open the Export Wizard, which lets you choose which
components to include:

| Component | Description |
|---|---|
| Data (CSV) | The current training dataset (including any injected outliers) |
| Fitted Models (RDS) | All currently fitted model objects as an R list |
| Plots (PNG / PDF) | Current boundary plots as image files |
| Grid (CSV) | The raw boundary prediction grid (x, y, prediction, probabilities) |
| Metrics (CSV) | The performance metrics table |
| Reproduce Script (R) | An auto-generated R script that reloads the data and redraws the plots |

The reproduce script loads `data.csv` and `models.rds` from the same folder and
calls `boundary_compute()` and `plot_boundary()` to regenerate the plots. It captures
the current grid resolution, zoom state, and visualization settings. UI-only state
(e.g., panel layout, color theme preferences) is intentionally not exported.

---

## Import Workspace Models

The **Import Workspace Models** panel scans your R Global Environment for `workflow`,
`model_fit`, and `model_spec` objects (from `tidymodels`). Any found objects can be
imported into the app for comparison against the built-in classifiers.
