---
title: "Getting Started with classbound"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with classbound}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The **classbound** package provides a streamlined interface for computing, visualizing, and interactively exploring classification decision boundaries in R. Whether you are teaching machine learning concepts, evaluating model performance, or debugging complex classifiers, `classbound` helps you understand how your models partition the feature space.

```{r setup}
library(classbound)
```

## Fitting a Model

`classbound` provides a unified `fit_model()` function that wraps various underlying algorithms. For example, to fit a decision tree:

```{r fit-model}
# Load the included example dataset
data(data69_1)

# Convert the target to a factor for classification
data69_1$Y <- as.factor(data69_1$Y)

# Fit an rpart decision tree model
model <- fit_model(
  data = data69_1[, c("V1", "V2")], 
  labels = data69_1$Y, 
  method = "rpart"
)
```

## Computing the Decision Boundary

Once a model is fitted, you can compute the boundary over a specified grid. Define the range for your features and the grid resolution:

```{r compute-boundary}
# Define the range for the grid
feature_range <- list(
  V1 = c(min(data69_1$V1), max(data69_1$V1)),
  V2 = c(min(data69_1$V2), max(data69_1$V2))
)

# Compute predictions over the grid
grid_data <- boundary_compute(
  model = model, 
  range = feature_range, 
  resolution = 100
)

head(grid_data)
```

## Visualizing the Boundary

To visualize the resulting boundary, pass the computed grid data to `plot_boundary()`. You can also overlay the original training observations to see how well the regions match the true classes.

```{r plot-boundary, fig.width=6, fig.height=4}
plot_boundary(
  boundary = grid_data, 
  obs_data = data69_1, 
  x_col = "V1", 
  y_col = "V2", 
  true_label = "Y"
)
```

## Interactive Exploration

For interactive exploration of data and models, you can launch the built-in Shiny application:

```r
explorapp()
```

This app allows you to upload data, select features, train models, and visualize boundaries on the fly.
