---
title: "Getting Started with classbound"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with classbound}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5
)
```

## What is classbound?

`classbound` is an R package for exploring and comparing classification decision boundaries. Given a fitted classifier and a dataset, it answers a simple question: *where in the feature space does the model change its prediction?*

The package supports:

- Any classifier whose `predict()` returns class labels
- 2D boundary plots for two-feature models
- High-dimensional data via 2D slicing or linear projection (PCA, tourr)
- Side-by-side multi-model comparison
- Probability surfaces (where the classifier provides probabilities)
- An interactive Shiny application (`explorapp()`) for visual exploration

## Installation

```{r eval=FALSE}
devtools::install_github("natydasilva/classbound")
```

## Quick start: one-step boundary plot

The `classbound()` wrapper fits a model, computes the decision boundary, and plots it in a single call.

```{r quickstart, message=FALSE, warning=FALSE, fig.width=9, fig.height=3}
library(classbound)
library(palmerpenguins)

penguins <- na.omit(penguins[, c("species", "bill_length_mm", "bill_depth_mm")])

classbound(
  data       = penguins,
  formula    = species ~ bill_length_mm + bill_depth_mm,
  classifier = rpart::rpart,
  resolution = 80
)
```

The colored regions show the predicted class for each point in the feature space. Training observations are overlaid as points colored by their true class.

## The modular API

For more control, use the three-step pipeline:

```{r pipeline, message=FALSE, warning=FALSE, fig.width=9, fig.height=3}
# Step 1: Fit the model
model <- fit_model(penguins, species ~ bill_length_mm + bill_depth_mm, rpart::rpart)

# Step 2: Compute the boundary grid
model <- boundary_compute(
  model,
  feature_range = list(bill_length_mm = c(30, 60), bill_depth_mm = c(10, 25)),
  resolution    = 80
)

# Step 3: Plot
plot_boundary(
  model,
  obs_data   = penguins,
  x_col      = "bill_length_mm",
  y_col      = "bill_depth_mm",
  true_label = "species"
)
```

`boundary_compute()` returns the same model object with `$boundary_data` populated. This means you can reuse the same fitted model with different visualizations, zoom levels, or color settings without refitting.

## Using different classifiers

`classbound` works with any classifier whose `predict()` method returns a vector of class labels. Most classifiers work automatically:

```{r classifiers, eval=FALSE}
# SVM (returns class labels natively; no extra work needed)
classbound(penguins, species ~ bill_length_mm + bill_depth_mm, e1071::svm)

# Random forest (matrix interface: randomForest expects x and y separately)
classbound(penguins, species ~ bill_length_mm + bill_depth_mm,
  randomForest::randomForest,
  interface = "matrix"
)
```

For classifiers whose `predict()` returns a list or other complex object, use the `predfun` argument to extract the class labels:

```{r predfun, eval=FALSE}
# MASS::qda returns a list, so extract $class manually
classbound(
  penguins,
  species ~ bill_length_mm + bill_depth_mm,
  MASS::qda,
  predfun = function(model, newdata, ...) predict(model, newdata, ...)$class
)
```

## Probability surface

Classifiers that return class probabilities (e.g., `rpart`, `randomForest`) enable a gradient visualization where decision regions are shaded by model confidence:

```{r gradient, message=FALSE, warning=FALSE, fig.width=9, fig.height=3}
model <- fit_model(penguins, species ~ bill_length_mm + bill_depth_mm, rpart::rpart)
model <- boundary_compute(model, resolution = 80)

plot_boundary(
  model,
  obs_data      = penguins,
  x_col         = "bill_length_mm",
  y_col         = "bill_depth_mm",
  true_label    = "species",
  show_gradient = TRUE
)
```

Deep colors indicate high model confidence; faded colors near boundaries indicate uncertainty.

Classifiers that do not provide probabilities (e.g., standard SVMs, PPtree) always show flat solid regions regardless of `show_gradient = TRUE`.

## Interactive exploration

For interactive exploration without writing code, launch the built-in Shiny application:

```{r explorapp, eval=FALSE}
# Launch with a dataset pre-loaded
explorapp(data = penguins, target_col = "species")

# Or launch empty and simulate data interactively
explorapp()
```

See the `vignette("explorapp-guide")` for a full walkthrough of the interactive features.

## What's next?

- **High-dimensional data**: `vignette("high-dimensional")` (2D slice vs. projection)  
- **Multiple models**: `vignette("tidymodels-workflow")` (comparing classifiers with `boundary_workflow_set()`)  
- **Custom classifiers**: `vignette("custom_adapters")` (`predfun` and custom S3 adapters)  
- **tourr integration**: `vignette("tourr-workflow")` (animated projection tours)  
