---
title: "Getting Started with classbound"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with classbound}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The **classbound** package provides a streamlined interface for computing, visualizing, and interactively exploring classification decision boundaries in R. Whether you are teaching machine learning concepts, evaluating model performance, or debugging complex classifiers, `classbound` helps you understand how your models partition the feature space.

```{r setup}
library(classbound)
```

## Fitting a Model

`classbound` provides a unified `fit_model()` function that wraps various underlying algorithms. For example, to fit a decision tree:

```{r message=FALSE, warning=FALSE}
# Load and prepare palmerpenguins dataset
library(palmerpenguins)
penguins <- na.omit(penguins[, -c(2, 7, 8)])

# Fit an rpart decision tree model predicting species
model <- fit_model(
  data = penguins[, c("bill_length_mm", "bill_depth_mm")],
  labels = penguins$species,
  method = "rpart"
)
```

## Computing the Decision Boundary

Once a model is fitted, you can compute the boundary over a specified grid. Define the range for your features and the grid resolution:

```{r}
# Define the range for the grid
feature_range <- list(
  bill_length_mm = c(30.0, 60.0),
  bill_depth_mm = c(10.0, 25.0)
)

# Compute predictions over the grid
grid_data <- boundary_compute(
  model = model,
  range = feature_range,
  resolution = 100
)

head(grid_data)
```

## Visualizing the Boundary

To visualize the resulting boundary, pass the computed grid data to `plot_boundary()`. You can also overlay the original training observations to see how well the regions match the true classes.

```{r}
plot_boundary(
  boundary = grid_data,
  obs_data = penguins,
  x_col = "bill_length_mm",
  y_col = "bill_depth_mm",
  true_label = "species"
)
```

## Interactive Exploration

For interactive exploration of data and models, you can launch the built-in Shiny application:

```{r eval=FALSE}
explorapp()
```

This app allows you to upload data, select features, train models, and visualize boundaries on the fly.
