---
title: "Comparing Classifiers with tidymodels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing Classifiers with tidymodels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 5
)
```

## Overview

`classbound` provides first-class support for the `tidymodels` ecosystem via
`boundary_workflow_set()`. Given a `workflow_set` of untrained or pre-trained classifiers,
it automatically fits each model, computes the decision boundary on a shared grid, and
returns a combined boundary data frame with a `model` column ready for faceted plotting.

## Required packages

```{r packages, eval=FALSE}
install.packages(c("tidymodels", "workflowsets", "parsnip", "rpart", "nnet"))
```

```{r load_pkgs, message=FALSE, warning=FALSE}
library(classbound)
library(palmerpenguins)
```

## Step 1: Prepare data

```{r data}
penguins <- na.omit(palmerpenguins::penguins[
  ,
  c("species", "bill_length_mm", "bill_depth_mm")
])
```

## Step 2: Define model specifications

Use `parsnip` to define model specifications independently of the fitting engine.

```{r specs, eval=requireNamespace("parsnip", quietly = TRUE) && requireNamespace("workflowsets", quietly = TRUE), message=FALSE}
library(parsnip)
library(workflowsets)

spec_tree <- decision_tree(mode = "classification") |>
  set_engine("rpart")

spec_rf <- rand_forest(mode = "classification") |>
  set_engine("randomForest")
```

## Step 3: Create a workflow set

A `workflow_set` pairs each model specification with a preprocessing formula.

```{r wf_set, eval=requireNamespace("parsnip", quietly = TRUE) && requireNamespace("workflowsets", quietly = TRUE)}
wf_set <- workflow_set(
  preproc = list(base = species ~ bill_length_mm + bill_depth_mm),
  models  = list(tree = spec_tree, forest = spec_rf)
)
wf_set
```

## Step 4: Compute boundaries for all models

`boundary_workflow_set()` handles fitting (if not already done) and boundary computation
for every workflow in the set. It returns a combined boundary data frame with a `model`
column identifying the `wflow_id`.

```{r compute, eval=requireNamespace("parsnip", quietly = TRUE) && requireNamespace("workflowsets", quietly = TRUE), message=FALSE, warning=FALSE}
bounds <- boundary_workflow_set(
  wf_set,
  data       = penguins,
  response   = "species",
  resolution = 60
)

# The result is a classbound object with multi-model boundary data
class(bounds)
head(bounds$boundary_data[, 1:4])
```

## Step 5: Plot with facets

`plot_boundary()` automatically facets multi-model objects by model name.

```{r plot, eval=requireNamespace("parsnip", quietly = TRUE) && requireNamespace("workflowsets", quietly = TRUE), message=FALSE, warning=FALSE}
plot_boundary(
  bounds,
  obs_data   = penguins,
  x_col      = "bill_length_mm",
  y_col      = "bill_depth_mm",
  true_label = "species"
)
```

## Step 6: Disagreement map

For two or more models, `type = "disagreement"` highlights where classifiers
predict differently — useful for identifying regions of high model uncertainty.

```{r disagree, eval=requireNamespace("parsnip", quietly = TRUE) && requireNamespace("workflowsets", quietly = TRUE), message=FALSE, warning=FALSE}
plot_boundary(bounds,
  type = "disagreement",
  x_col = "bill_length_mm", y_col = "bill_depth_mm"
)
```

## Using pre-trained workflows

If your workflows are already trained (e.g., from `tune::fit_resamples()` or a
previous call to `parsnip::fit()`), `boundary_workflow_set()` detects this and
skips refitting.

```{r pretrained, eval=FALSE}
# Fit individually first
wf1 <- workflows::workflow(species ~ ., spec_tree) |> parsnip::fit(penguins)
wf2 <- workflows::workflow(species ~ ., spec_rf) |> parsnip::fit(penguins)

# Wrap in a workflow_set (already trained)
wf_trained <- workflowsets::workflow_set(
  preproc = list(base = species ~ .),
  models  = list(tree = spec_tree, forest = spec_rf)
)
# boundary_workflow_set() will refit because wf_set workflows are not trained
# Use as_classbound() directly for pre-fitted objects:
m1 <- as_classbound(wf1, data = penguins, response = "species")
m2 <- as_classbound(wf2, data = penguins, response = "species")
bounds_manual <- boundary_compute(
  list(tree = m1, forest = m2),
  feature_range = list(bill_length_mm = c(30, 60), bill_depth_mm = c(10, 25)),
  resolution = 60
)
plot_boundary(bounds_manual,
  obs_data = penguins,
  x_col = "bill_length_mm", y_col = "bill_depth_mm",
  true_label = "species"
)
```
